Behind the Delays: Exploring Variations in Toronto’s Transit Disruptions

### 2024-11-27 paper/paper.qmd
citation upload and structure updated

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

The remainder of this paper is structured as follows. @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy. @sec-models outlines our modeling approach..., incorporating lessons learned from recent electoral cycles. Our predictions are under section of each model. @sec-discuss discusses the implications of our findings and suggests directions for future research. Finally, @sec-appendix evaluates methodology and survey copy.

@@

Following @tellingstories, we consider...

@@

# Model

@@

Define $y_i$ as the duration of delay (in seconds) for a given transit event. Let $\beta_i$ represent the transit mode, encoded as an indicator variable for subway, streetcar, or bus, and $\gamma_i$ represent the time of day, categorized into morning peak, midday, evening peak, and late night.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

# Discussion

@@

# Appendix {-}

>>>>

The remainder of this paper is structured as follows. @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy. @sec-model outlines our modeling approach..., incorporating lessons learned from recent electoral cycles. Our predictions are under section of each model. @sec-discuss discusses the implications of our findings and suggests directions for future research. Finally, @sec-appendix evaluates methodology and survey copy.

@@

@@

# Model {#sec-model}

@@

## Model Set-Up

Define $y_i$ as the duration of the delay (in minutes) for a given transit event. Let $\beta_i$ represent the transit mode (subway, streetcar, or bus), and $\gamma_i$ and $\delta_i$ represent the time and day of the week, respectively. The generative model is specified as follows:

\begin{align} 
y_i | \mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i + \delta_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\delta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Here:
\begin{itemize}
    \item $y_i$ is the observed delay duration for a specific transit event.
    \item $\alpha$ is the baseline average delay duration.
    \item $\beta_i$ captures the effect of transit mode (subway, streetcar, or bus).
    \item $\gamma_i$ captures the effect of the time of day (e.g., peak or off-peak).
    \item $\delta_i$ captures the effect of the day of the week.
    \item $\sigma$ is the standard deviation, representing unexplained variability in $y_i$.
\end{itemize}

This Bayesian model allows for estimating the contributions of transit mode, time, and day to delay durations while accounting for uncertainty in the data.

@@

# Discussion {#sec-discuss}

@@

# Appendix {#sec-appendix}
